---
title: "classwork"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(primer.data)
library(rstanarm)
library(tidyverse)
```

# Causal vs. Predictive Models

you can use a causal model to make predictions, but it can do more. 

income ~ liberalism

- Predictive models care nothing about causation
- Every model is predictive, in the sense that, if we give you new data — and it is drawn from a stable distribution — then you can create a predictive forecast
- so it has some quality that cannot be changed/ is independent of new factor--> for example age
ex. if you change your political ideology your pay won't change

Causal->
- existing correlative relationship?
- "no causation without manipulation" - you need some type of treatment that can manipulate the situation
- potential outcomes- change in income
- cannot randomly assign some people to be liberal or non- liberal

potential outcomes of what?
  - income as outcome variable-> income over 100k and below 100k

```{r}
# Today, we will look at liberal as a function of income. Note that we should
# first transform income by dividing it by 100,000. That is, income is now in
# units of 100k dollars, whereas before it was in units of one dollar. This does
# not change the substance, but it makes interpretation easier.

#Today’s question: Consider the set of people who make $100,000. What proportion of them are liberal?
  
# Create a model, fit_1, of liberal as a function of income. Use this model to
# answer our question. Create a nice graphic.

new_trains <- trains %>% 
                mutate(small_income = income/100000)

fit_1 <-  <- stan_glm(data = new_trains, 
                            formula = liberal ~ small_income, 
                            family = binomial,
                            refresh = 0,
                            seed = 87)

print(fit_1)
```

y_1 = beta_1 + beta_0 + income_small + epsilon_1
```{r notes}
# family relates to the left hand side variable 0/1 binomial, gaussian- continuous 
#mad_sd- > spread
#median- > location
# median of our posterior distribution for mu (beta_0) -> the median .6, 
# 95% confident that 1.0 between .2 mean 
#mad_sd standard error -> .4


newobs <- tibble(small_income = 1)

pe <- posterior_epred(fit_1, newobs) %>% 
    as_tibble() 
    
pe %>% 
   ggplot(aes(x = `1`)) +
    geom_histogram(aes(y = after_stat(count/sum(count))),
                   bins = 100) +
    labs(title = "Posterior Probability Distribution for the Percentake of Population
         Which Makes $100,000",
         subtitle = "The population of people that make 100k is likely to have 
         a similar number of liberal and non-liberal people",
         x = "Probability that Population is Liberal ",
         y = "Probability",
         caption = "Data source: trains") + 
    scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    theme_classic()

```

